Dynamic Witnesses for Static Type Errors

### 2015-11-20 paper/pldi16-submission/introduction.tex
done with intro

% This process of iterative expansion of the trace ensures that
% the context of intermediate steps is not lost; whereas traditional
% debuggers force one to move between statements, in our visualization
% you never lose sight of where you came from.

@@

\paragraph{3. Evaluating Witnesses}
%
Our third contribution is an extensive evaluation of our approach
on two different sets of ill-typed programs obtained by instrumenting
compilers used in beginner's classes~(\S~\ref{sec:eval}).
%
The first is the \uwbench\ data set~\cite{lerner_searching_2007}
comprising \uwsize\ ill-typed programs.
%
The second is a new \ucsdbench\ data set, comprising \ucsdsize\
ill-typed programs.
%
We show that for both data sets, our technique is able to generate
witnesses for nearly 90\% of the programs, in under a second the
vast majority of cases.
%
Furthermore, we show that a simple interactive strategy yields
compact counterexample traces with fewer than 5 steps for 60\%
of the programs, and under 10 steps for 90\% of the programs.

In short, our results show that at least for beginners, ill-typed
programs \emph{do} go wrong, in the vast majority of cases, which,
in turn, opens the door to a completely different, dynamic way
to explain, understand, and appreciate the benefits of static typing.

%% Then, we formalize our algorithm and prove it sound
%% (\S~\ref{sec:searching-witness}).
%% %
%% Next, we extend the algorithm to collect a reduction graph and formalize
%% a set of traversals over the graph (\S~\ref{sec:explaining}).
%% %
%% Finally, we evaluate our search algorithm and trace generation on a
%% large set of real student programs (\S~\ref{sec:evaluation}).
%%
%% \paragraph{Outline}
%% We begin the rest of the paper with an overview of our search algorithm
%% and trace visualization (\S~\ref{sec:overview}).

% Then, we formalize our algorithm and prove it sound
% (\S~\ref{sec:searching-witness}), and evaluate it on a large set of real
% student programs (\S~\ref{sec:eval}).
% %
% Next, we extend the search algorithm to collect the reduction graph and
% formalize a set of traversals (\S~\ref{sec:explaining}), and measure the
% complexity of these traces on our benchmarks
% (\S~\ref{sec:evaluation-nanomaly}).
%
% Finally, we present the results of a preliminary user study on the full
% system (\S~\ref{sec:user-study}).

>>>>

@@

\paragraph{3. Evaluating Witnesses}
%
Our third contribution is an extensive evaluation of our approach
on two different sets of ill-typed programs obtained by instrumenting
compilers used in beginner's classes~(\S~\ref{sec:evaluation}).
%
The first is the \uwbench\ data set~\cite{lerner_searching_2007}
comprising \uwsize\ ill-typed programs.
%
The second is a new \ucsdbench\ data set, comprising \ucsdsize\
ill-typed programs.
%
We show that for both data sets, our technique is able to generate
witnesses for nearly 90\% of the programs, in under a second the
vast majority of cases.
%
Furthermore, we show that a simple interactive strategy yields
compact counterexample traces with fewer than 5 steps for 60\%
of the programs, and under 10 steps for 90\% of the programs.

\smallskip
In short, our results show that at least for beginners, ill-typed
programs \emph{do} go wrong, in the vast majority of cases, which,
in turn, opens the door to a novel dynamic way to explain, understand,
and appreciate the benefits of static typing.
